---
title: "Crossroads Baseball Project."
author: "Levi Cain, Kara Godsey, and Dr. Ryan Johnson"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    #highlight: kate
---


```{r setup, echo=FALSE, cache=FALSE}
suppressWarnings(library(knitr))
suppressWarnings(library(rmdformats))
suppressWarnings(library(kableExtra))

## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```


# Welcome
Welcome to our student research project.  This project is still in its beginning stages.  The current plan is to rank players and predict games.

### Update 2/12/2020
We have scraped data from the website, although the strictly-conference level data is still elluding us.  To view our process go to "Scraping the Data" on the left-hand side of your screen.

---

# Scraping the Data

For this project we used the NAIA statistics posted on the website [dakstats](http://www.dakstats.com/websync/Pages/Association.aspx?association=10).  They have batting, pitching, and fielding statitistics for each baseball team in the Crossroads league, of which Grace College is a member.

We found the `rvest` R package to be very useful for this project.

```{r}
suppressWarnings(library(rvest))
```

## The Webpage Url's

Dakstats assigns each team an identification number, and uses this number in their webpage urls.  We've created 2-4 letter abreviations for each team, and named them by their ID numbers.

```{r teams-and-urls}
teams <- c("BC","GOC","GRC","HU","INWU","MAR","MVNU","SAU","SFIN","TAYL")
names(teams) <- c(1629, 1678, 1679, 1688, 1694, 1717, 1736, 1780, 1805, 1784)
urls <- paste0("http://www.dakstats.com/WebSync/Pages/Team/IndividualStats.aspx?association=10&sg=MBA&conference=NAIMBA1_CROSS&team=",names(teams),"&sea=NAIMBA_2019")
```
## Scraping the tables using the `rvest` package

Then we created functions to retrieve only the batting, pitching, and fielding tables, which we turned into data frames in R.

```{r retrieval-functions}
get_batting <- function(url) {
  bpf <- url %>%
    read_html() %>%
    html_nodes("table") %>%
    .[[37]] %>%
    html_table(fill = TRUE)
}

get_pitching <- function(url) {
  bpf <- url %>%
    read_html() %>%
    html_nodes("table") %>%
    .[[38]] %>%
    html_table(fill = TRUE)
}

get_fielding <- function(url) {
  bpf <- url %>%
    read_html() %>%
    html_nodes("table") %>%
    .[[39]] %>%
    html_table(fill = TRUE)
}
```

We used the `lapply` R function with these functions to get our data.

```{r lapply-it}
batting <- lapply(urls, get_batting)
names(batting) <- teams
pitching <- lapply(urls, get_pitching)
names(pitching) <- teams
fielding <- lapply(urls, get_fielding)
names(fielding) <- teams
```

## Saving the data

And put all of our data into one list.

```{r}
baseball <- list(batting, pitching, fielding)
names(baseball) <- c("Batting", "Pitching", "Fielding")
```
Then we wrote a function to write all of the tables into csv files,

```{r writeit}
#A function to write the data to csv files
writeit <- function(x) {
  #Batting
  write.csv(batting[[x]], file = paste0("Data/Batting", "/", x, ".csv"))
  #Pitching
  write.csv(pitching[[x]], file = paste0("Data/Pitching", "/", x, ".csv"))
  #Fielding
  write.csv(fielding[[x]], file = paste0("Data/Fielding", "/", x, ".csv"))
}
```
and used `lapply` to write the csv's all at once.

```{r write-to-csv}
invisible(lapply(teams, writeit))
```
Lastly, we also saved our data to an Rdata file for our own convenience.

```{r saved}
save(baseball, file = "Data/stats_by_type.Rdata")
```

---

# Creating the Park Factor Code

We wanted to be able to adjust our statistics based on the ballpark where the game was played or will be played. We scraped the schedule data from [dakstats](https://www.dakstats.com/WebSync/Pages/Team/TeamSchedule.aspx?association=10&sg=MBA&sea=NAIMBA_2019&team=1679)
using code similar to the code found in Scraping the Data.

```{r schedule_scrape}
library(rvest)


teams <- c("BC","GOC","GRC","HU","INWU","MAR","MVNU","SAU","SFIN","TAYL")
#R lets us rename the indices of vectors.
names(teams) <- c(1629, 1678, 1679, 1688, 1694, 1717, 1736, 1780, 1805, 1784)
  p_effects <- paste0("http://www.dakstats.com/WebSync/Pages/Team/TeamSchedule.aspx?association=10&sg=MBA&sea=NAIMBA_2019&team=", names(teams))
  
  #Function to get schedule of each team and remove blank rows
  
  get_schedule <- function(url){
  tbls_effects <- url %>%
    read_html() %>%
    html_nodes("table") %>%
    .[36] %>%
    html_table(fill = TRUE)
  Team_Schedule <- data.frame(tbls_effects[[1]])
  Team_Schedule <- Team_Schedule[!apply(is.na(Team_Schedule) | Team_Schedule == "", 1, all),]
  }
  
  schedule <- lapply(p_effects, get_schedule)
  names(schedule) <- teams
```

## Park Factor Function

Since we were looking specifically at Crossroads League conference statistics, we wanted to calculate park factor using only conference games. Therefore, we created a new dataframe with only conference games.

Next, because park factor ajdusts based on runs scored at home and runs scored away, we separated home and away games. 

We then put the scores specific to home or away games into a vector.

Finally, we used the lapply to apply the function to each team.
```{r Park_Factor function}
  Park_effects <- function(g){
    
    Team_Schedule <- schedule[[g]]
    Sched <- Team_Schedule["X2"]
    Scores <- c()
    Conf_Games <- data.frame(Date=character(),
                             Opponent=character(), 
                             Location=character(),
                             Score=character(),
                             Result=character(),
                             stringsAsFactors=FALSE) 
    j=1
    for (i in 1:nrow(Sched)){
    
      if (grepl("*", Sched[[1]][[i]], fixed = TRUE) == TRUE){
        Conf_Games[j,] <- Team_Schedule[i,]
        j = j + 1
      } 
    }
    
    #print(Conf_Games)
    
    #Separate Home games from neutral and away games
    Home_Games <- data.frame(Date=character(),
                             Opponent=character(), 
                             Location=character(),
                             Score=character(),
                             Result=character(),
                             stringsAsFactors=FALSE)
    
    Away_Games <- data.frame(Date=character(),
                             Opponent=character(), 
                             Location=character(),
                             Score=character(),
                             Result=character(),
                             stringsAsFactors=FALSE)
    
    
    Loc <- Conf_Games["Location"]
  
    w=1
    z=1
    for (k in 1:nrow(Loc)){
      if (is.element("H", Loc[[1]][[k]]) == TRUE){
        Home_Games[w,] <- Conf_Games[k,]
        w = w + 1
      }
      else if (is.element("A", Loc[[1]][[k]]) == TRUE){
        Away_Games[z,] <- Conf_Games[k,]
        z = z + 1
      }
     
    }
    
    #Put Home team score for each game into a vector
      
  
      Scores_H <- Home_Games[["Score"]]
      Scores_A <- Away_Games[["Score"]]
      p=1
      q=1
      c=1
      d=1
      
        Team_Runs_A <- c()
        Team_Runs_H <- c()
        Opp_Runs_H <- c()
        Opp_Runs_A <- c()
        
        for (x in Scores_H){
          
          if (grepl("(", x, fixed = TRUE) == TRUE){
            Opp_Runs_H[d] <- as.integer(substr(x, (regexpr(pattern ='-', x)+1), nchar(x)-4))
            Team_Runs_H[p] <- as.integer(substr(x, 1, (regexpr(pattern ='-', x)-1)))
          }
          else{
            Team_Runs_H[p] <- as.integer(substr(x, 1, (regexpr(pattern ='-', x)-1)))
            Opp_Runs_H[d] <- as.integer(substr(x, (regexpr(pattern ='-', x)+1), nchar(x)))
            
          }
          
          p = p + 1
          d = d + 1
        }
        
        for (r in Scores_A){
          if (grepl("(", r, fixed = TRUE) == TRUE){
            
            Opp_Runs_A[c] <- as.integer(substr(r, (regexpr(pattern ='-', r)+1), nchar(r)-4))
            Team_Runs_A[q] <- as.integer(substr(r, 1, (regexpr(pattern ='-', r)-1)))
            
          }
          else{
            
            Team_Runs_A[q] <- as.integer(substr(r, 1, (regexpr(pattern ='-', r)-1)))
            Opp_Runs_A[c] <- as.integer(substr(r, (regexpr(pattern ='-', r)+1), nchar(r)))
            
          }
          q = q + 1
          c = c + 1
          
        }
       
  
  Park_effect <- ((sum(Team_Runs_H)+sum(Opp_Runs_H))/length(Team_Runs_H))/((sum(Team_Runs_A)+sum(Opp_Runs_A))/length(Team_Runs_A))
      
  }
  
  p_factor <- teams %>%
    lapply(Park_effects)
  names(p_factor) <- teams
```

---

# Batting Statistics

## The Basics

First, we loaded the data tables that we previously created.

```{r}
load("Data/stats_by_type.Rdata")
Batting <- baseball[["Batting"]]
teams <- names(Batting)
```

Next, we created a function to calculate advanced statistics using the base data in the tables. The statistics we chose to add to our tables include: Caught Stealing, Runs Created (Basic), OPS, POP, Total Average, Batting Average on Balls in Play (BABIP), and the Hoban Efficiency Quotient for offense.

```{r}
bat_stats <- function(y){
  x <- Batting[[y]]
  #Caught Stealing
  x$CS <- x$SBA - x$SB
  #Singles
  x$X1B <- x$H - x[,8] - x[,9] - x$HR
  #Runs Created
  x$RC <- round((x$H + x$BB)*x$TB/(x$AB + x$BB),2)
  #OPS
  x$OPS <- x$OB + x$SLG
  #POP
  x$POP <- x$OPS + x$AVG
  #Total Average
  x$TA <- round((x$TB + x$BB + x$HBP + x$SB)/(x$AB - x$H + x$SH + x$SF + x$CS + x$GDP), 3)
  #Plate Appearances
  x$PA <- x$AB+x$BB+x$HBP+x$SF
  #Batting Average on Balls in Play
  x$BABIP <- round((x$H - x$HR)/(x$AB - x$SO - x$HR + x$SF),3)
  #Hoban Efficiency Quotient - offense
  x$HEQO <- x$TB + x$R + x$RBI + x$SB + (0.5*x$BB)
  #Base Stealing Runs
  x$Steal_Runs <- round(((x$H + x$BB - x$CS)*(x$TB + (0.7 * x$SB)))/(x$AB + x$BB + x$CS),2)
  x
}
```

Then we used the `lapply` R function to apply these statistics to each of the teams, resulting in advanced statistics for every player in the league.

```{r lapplyit}
Batting <- teams %>%
  lapply(bat_stats)
names(Batting) <- teams
```

## Crossroads League

Next, we calculated league totals for the basic statistics, for the purpose of calculating more advanced individual statistics later on.

```{r}
batting_stats <- c("H","AB","BB","HBP","SF","TB", "2B","3B","HR","X1B","R")
CL_Avg_B <- c()
All_stats_batting <- c()
w = 0

for (n in batting_stats){
  
  for (j in teams){
    
    P <- Batting[[j]]#as.data.frame(read.csv(paste("Data/Batting/",j,".csv",sep="")))
    
    col_num =  which( colnames(P)==n )
    for (k in seq(length(P[[n]]))){ 
      
      All_stats_batting[w+k] <- P[,col_num][k]
      #print(All_stats)
    }
    
    w  = k + w
    
  }
  CL_Avg_B[n] <- sum(All_stats_batting)
  All_stats_batting <- c()
  w = 0
}

print(CL_Avg_B)
```

Then, we calculated advanced statistics for the Crossroads League as a whole. These included: On-Base Percentage (OBP), Slugging Percentage (SLG), weighted On-Base Percentage (wOBA), Plate Appearances (PA), Runs per Plate Appearance (R/PA), weighted Runs Created (wRC), and weighted Runs Created per Plate Appearance (wRC/PA). Additionally, we calculated ABF, which is a constant that is necessary to caluclate batting runs later on.

```{r}
CL_OBP <- (CL_Avg_B[[1]]+CL_Avg_B[[3]]+CL_Avg_B[[4]])/(CL_Avg_B[[2]]+CL_Avg_B[[3]]+CL_Avg_B[[4]]+CL_Avg_B[[5]])
CL_SLG <- (CL_Avg_B[[6]])/(CL_Avg_B[[2]])
CL_wOBA <- ((.69)*(CL_Avg_B[["BB"]])+(.72)*(CL_Avg_B[["HBP"]])+(.89)*(CL_Avg_B[["X1B"]])+(1.27)*(CL_Avg_B[["2B"]])+(1.62)*(CL_Avg_B[["3B"]])+(2.10)*(CL_Avg_B[["HR"]]))/(CL_Avg_B[["AB"]]+CL_Avg_B[["BB"]]+CL_Avg_B[["SF"]]+CL_Avg_B[["HBP"]])
CL_PA <- CL_Avg_B[["AB"]]+CL_Avg_B[["BB"]]+CL_Avg_B[["HBP"]]+CL_Avg_B[["SF"]] 
CL_R_per_PA <- CL_Avg_B[["R"]]/CL_PA
CL_wRC <- CL_Avg_B[["R"]]
CL_wRC_per_PA <- CL_R_per_PA

#League adjustment for Batting Runs
ABF <- ((0.47 * CL_Avg_B[["X1B"]]) + (0.38*CL_Avg_B[["2B"]]) + (0.55*CL_Avg_B[["3B"]]) + (0.93*CL_Avg_B[["HR"]]) +(0.33*(CL_Avg_B[["BB"]]+CL_Avg_B[["HBP"]]))/(CL_Avg_B[["AB"]]-CL_Avg_B[["H"]]))

```

## Advanced Statistics

Now that we have these league statistics, we can add more advanced statistics to our tables. The statistics we added were: OPS+, wOBA, Batting Runs, weighted Runs Above Average (wRAA), weighted Runs Created (wRC), and weighted Runs Created plus (wRC+). All of these statistics adjust the individual player's performance to reflect their contribution in comparison to other players in the league.  

```{r}
bat_stats2 <- function(y){
  x <- Batting[[y]]
  #OPS+
  x$OPS_plus <- round((100)*(((x$OB/CL_OBP)+(x$SLG/CL_SLG))-1),2)
  #weighted On-Base Average
  x$wOBA <- round(((.69)*(x$BB)+(.72)*(x$HBP)+(.89)*(x$X1B)+(1.27)*(x[["2B"]])+(1.62)*(x[["3B"]])+(2.10)*(x$HR))/(x$AB+x$BB+x$SF+x$HBP),3)
  #Batting Runs
  x$Bat_Runs <- round(((0.47*x$X1B) + (0.38*x[["2B"]]) + (0.55*x[["3B"]]) + (0.93*x$HR) + ((0.33)*x$BB + x$HBP)/(ABF *(x$AB - x$H))),2)
  #weighted Runs Above Average
  x$wRAA <- round(((x$wOBA-CL_wOBA)/(1.157))*(x$PA),3)
  #weighted Runs Created
  x$wRC <- round(x$wRAA + (x$PA)*((CL_Avg_B[["R"]])/(CL_PA)),2)
  x$wRC_plus <- round((100)*((((x$wRAA/x$PA)+CL_R_per_PA)+(CL_R_per_PA-(p_factor[[y]])*(CL_R_per_PA)))/(CL_wRC_per_PA)),2)
  x
}
```

Finally, we used the `lapply` R function again to apply these statistics to all the teams in the Crossroads League.

```{r lapply}
Batting <- teams %>%
  lapply(bat_stats2)
names(Batting) <- teams
```

Here is an example of the resulting tables that we produced.

```{r echo=FALSE}
kable(Batting[["GRC"]]) %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "condensed"), full_width = FALSE)
```

---