---
title: "Crossroads Baseball Project."
author: "Levi Cain, Kara Godsey, and Dr. Ryan Johnson"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    #highlight: kate
---


```{r setup, echo=FALSE, cache=FALSE}
suppressWarnings(library(knitr))
suppressWarnings(library(rmdformats))
suppressWarnings(library(kableExtra))
suppressWarnings(library(ggplot2))
suppressWarnings(library(DT))
suppressWarnings(library(ggrepel))

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```


# Welcome! 

These are the results from our project.  You can check out how we did everything here: <li><a href="https://mathatgrace.github.io/CrossRoads-Baseball/">Project</a></li>


## Overview

On this site you will find park factors, the top hitters in the conference, projections for upcoming games, and a searchable table containing stats for all hitters in the Crossroads League.

---

---

```{r}
suppressWarnings(library(rvest))
```


```{r teams-and-urls}
teams <- c("BC","GOC","GRC","HU","INWU","MAR","MVNU","SAU","SFIN","TAYL")
names(teams) <- c(1629, 1678, 1679, 1688, 1694, 1717, 1736, 1780, 1805, 1784)
urls <- paste0("http://www.dakstats.com/WebSync/Pages/Team/IndividualStats.aspx?association=10&sg=MBA&conference=NAIMBA1_CROSS&team=",names(teams),"&sea=NAIMBA_2019")
```

```{r retrieval-functions}
get_batting <- function(url) {
  bpf <- url %>%
    read_html() %>%
    html_nodes("table") %>%
    .[[37]] %>%
    html_table(fill = TRUE)
}

get_pitching <- function(url) {
  bpf <- url %>%
    read_html() %>%
    html_nodes("table") %>%
    .[[38]] %>%
    html_table(fill = TRUE)
}

get_fielding <- function(url) {
  bpf <- url %>%
    read_html() %>%
    html_nodes("table") %>%
    .[[39]] %>%
    html_table(fill = TRUE)
}
```


```{r lapply-it}
batting <- lapply(urls, get_batting)
names(batting) <- teams
pitching <- lapply(urls, get_pitching)
names(pitching) <- teams
fielding <- lapply(urls, get_fielding)
names(fielding) <- teams
```

```{r}
baseball <- list(batting, pitching, fielding)
names(baseball) <- c("Batting", "Pitching", "Fielding")
```

```{r writeit}
#A function to write the data to csv files
writeit <- function(x) {
  #Batting
  write.csv(batting[[x]], file = paste0("Data/Batting", "/", x, ".csv"))
  #Pitching
  write.csv(pitching[[x]], file = paste0("Data/Pitching", "/", x, ".csv"))
  #Fielding
  write.csv(fielding[[x]], file = paste0("Data/Fielding", "/", x, ".csv"))
}
```

```{r write-to-csv}
invisible(lapply(teams, writeit))
```

```{r saved}
save(baseball, file = "Data/stats_by_type.Rdata")
```


```{r schedule_scrape}
library(rvest)


teams <- c("BC","GOC","GRC","HU","INWU","MAR","MVNU","SAU","SFIN","TAYL")
#R lets us rename the indices of vectors.
names(teams) <- c(1629, 1678, 1679, 1688, 1694, 1717, 1736, 1780, 1805, 1784)
  p_effects <- paste0("http://www.dakstats.com/WebSync/Pages/Team/TeamSchedule.aspx?association=10&sg=MBA&sea=NAIMBA_2019&team=", names(teams))
  
  #Function to get schedule of each team and remove blank rows
  
  get_schedule <- function(url){
  tbls_effects <- url %>%
    read_html() %>%
    html_nodes("table") %>%
    .[36] %>%
    html_table(fill = TRUE)
  Team_Schedule <- data.frame(tbls_effects[[1]])
  Team_Schedule <- Team_Schedule[!apply(is.na(Team_Schedule) | Team_Schedule == "", 1, all),]
  }
  
  schedule <- lapply(p_effects, get_schedule)
  names(schedule) <- teams
```

```{r Park_Factor function}
  Park_effects <- function(g){
    
    Team_Schedule <- schedule[[g]]
    Sched <- Team_Schedule["X2"]
    Scores <- c()
    Conf_Games <- data.frame(Date=character(),
                             Opponent=character(), 
                             Location=character(),
                             Score=character(),
                             Result=character(),
                             stringsAsFactors=FALSE) 
    j=1
    for (i in 1:nrow(Sched)){
    
      if (grepl("*", Sched[[1]][[i]], fixed = TRUE) == TRUE){
        Conf_Games[j,] <- Team_Schedule[i,]
        j = j + 1
      } 
    }
    
    #print(Conf_Games)
    
    #Separate Home games from neutral and away games
    Home_Games <- data.frame(Date=character(),
                             Opponent=character(), 
                             Location=character(),
                             Score=character(),
                             Result=character(),
                             stringsAsFactors=FALSE)
    
    Away_Games <- data.frame(Date=character(),
                             Opponent=character(), 
                             Location=character(),
                             Score=character(),
                             Result=character(),
                             stringsAsFactors=FALSE)
    
    
    Loc <- Conf_Games["Location"]
  
    w=1
    z=1
    for (k in 1:nrow(Loc)){
      if (is.element("H", Loc[[1]][[k]]) == TRUE){
        Home_Games[w,] <- Conf_Games[k,]
        w = w + 1
      }
      else if (is.element("A", Loc[[1]][[k]]) == TRUE){
        Away_Games[z,] <- Conf_Games[k,]
        z = z + 1
      }
     
    }
    
    #Put Home team score for each game into a vector
      
  
      Scores_H <- Home_Games[["Score"]]
      Scores_A <- Away_Games[["Score"]]
      p=1
      q=1
      c=1
      d=1
      
        Team_Runs_A <- c()
        Team_Runs_H <- c()
        Opp_Runs_H <- c()
        Opp_Runs_A <- c()
        
        for (x in Scores_H){
          
          if (grepl("(", x, fixed = TRUE) == TRUE){
            Opp_Runs_H[d] <- as.integer(substr(x, (regexpr(pattern ='-', x)+1), nchar(x)-4))
            Team_Runs_H[p] <- as.integer(substr(x, 1, (regexpr(pattern ='-', x)-1)))
          }
          else{
            Team_Runs_H[p] <- as.integer(substr(x, 1, (regexpr(pattern ='-', x)-1)))
            Opp_Runs_H[d] <- as.integer(substr(x, (regexpr(pattern ='-', x)+1), nchar(x)))
            
          }
          
          p = p + 1
          d = d + 1
        }
        
        for (r in Scores_A){
          if (grepl("(", r, fixed = TRUE) == TRUE){
            
            Opp_Runs_A[c] <- as.integer(substr(r, (regexpr(pattern ='-', r)+1), nchar(r)-4))
            Team_Runs_A[q] <- as.integer(substr(r, 1, (regexpr(pattern ='-', r)-1)))
            
          }
          else{
            
            Team_Runs_A[q] <- as.integer(substr(r, 1, (regexpr(pattern ='-', r)-1)))
            Opp_Runs_A[c] <- as.integer(substr(r, (regexpr(pattern ='-', r)+1), nchar(r)))
            
          }
          q = q + 1
          c = c + 1
          
        }
       
  
  Park_effect <- ((sum(Team_Runs_H)+sum(Opp_Runs_H))/length(Team_Runs_H))/((sum(Team_Runs_A)+sum(Opp_Runs_A))/length(Team_Runs_A))
      
  }
  
  p_factor <- teams %>%
    lapply(Park_effects)
  names(p_factor) <- teams
```

# Park Factor

Park Factor is a measurement of an individual ballpark's potential impact on the game. This number will take into account factors such as altitude and dimensions of the field. In the graph below, the line segment in the middle represents a park value of 1, which is considered a "neutral" park. 

A value greater than 1 indicates the park favors the hitter. Oftentimes this suggests more total hits and/or more total runs will be scored than an average game played at a neutral park. For example, in Major League Baseball, the home field for the Colorado Rockies, Coors Field, had a park factor of 1.33 in 2019. Coors Field is located in Denver, meaning the park sits one mile above sea level. At this altitude a baseball carries farther, resulting in more home runs. Thus, Coors Field is a hitter friendly park. Altitude is not the only cause of differences in park factors. Small parks, such as Fenway Park in Boston, also are considered hitter friendly. The fences at Fenway are closer to home plate than at most other parks. Thus, a baseball hit 340 feet to right field may be a home run in Boston, while being an out in Milwaukee. This is the significance of the park factor statistic. 

```{r park factor plot}
pf <- c(0.964, 1.12, 0.887, 0.732, 1.04, 0.794, 0.857, 1.15, 1.35, 1.34)
park <- data.frame(p_factor)
tpark <- data.frame(t(park))

ggplot(tpark, aes(x=pf, y=0)) +
  geom_point(size = 8) + 
  annotate("segment", x = 0.6, xend = 1.4, y=0, yend=0, size = 1) +
  annotate("segment", x=0.6,xend=0.6,y=-.1,yend=0.1,size=1) +
  annotate("text",x=0.65,y=.5,size =4,label = "Pitcher") +
  annotate("text",x=0.65,y=.4,size=4,label = "friendly") +
  annotate("segment",x=1, xend = 1, y=-0.1, yend=0.1, size = 1) +
  annotate("segment", x=1.4, xend=1.4, y=-0.1, yend=0.1, size=1) + 
  annotate("text", x=1.35, y=.5, size=4, label = "Hitter") +
  annotate("text",x=1.35, y=.4, size=4, label = "friendly") +
  geom_text_repel(aes(label = teams), col="black", vjust = 3) +
  scale_x_continuous(limits = c(.6, 1.4)) + 
  scale_y_continuous(limits = c(-1,1)) + 
  theme(panel.background = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), axis.title = element_blank())

```


```{r}
load("Data/stats_by_type.Rdata")
Batting <- baseball[["Batting"]]
teams <- names(Batting)
```


```{r}
bat_stats <- function(y){
  x <- Batting[[y]]
  #Caught Stealing
  x$CS <- x$SBA - x$SB
  #Singles
  x$X1B <- x$H - x[,8] - x[,9] - x$HR
  #Runs Created
  x$RC <- round((x$H + x$BB)*x$TB/(x$AB + x$BB),2)
  #OPS
  x$OPS <- x$OB + x$SLG
  #POP
  x$POP <- x$OPS + x$AVG
  #Total Average
  x$TA <- round((x$TB + x$BB + x$HBP + x$SB)/(x$AB - x$H + x$SH + x$SF + x$CS + x$GDP), 3)
  #Plate Appearances
  x$PA <- x$AB+x$BB+x$HBP+x$SF
  #Batting Average on Balls in Play
  x$BABIP <- round((x$H - x$HR)/(x$AB - x$SO - x$HR + x$SF),3)
  #Hoban Efficiency Quotient - offense
  x$HEQO <- x$TB + x$R + x$RBI + x$SB + (0.5*x$BB)
  #Base Stealing Runs
  x$Steal_Runs <- round(((x$H + x$BB - x$CS)*(x$TB + (0.7 * x$SB)))/(x$AB + x$BB + x$CS),2)
  x
}
```

```{r lapplyit}
Batting <- teams %>%
  lapply(bat_stats)
names(Batting) <- teams
```

```{r}
batting_stats <- c("H","AB","BB","HBP","SF","TB", "2B","3B","HR","X1B","R")
CL_Avg_B <- c()
All_stats_batting <- c()
w = 0

for (n in batting_stats){
  
  for (j in teams){
    
    P <- Batting[[j]]#as.data.frame(read.csv(paste("Data/Batting/",j,".csv",sep="")))
    
    col_num =  which( colnames(P)==n )
    for (k in seq(length(P[[n]]))){ 
      
      All_stats_batting[w+k] <- P[,col_num][k]
      #print(All_stats)
    }
    
    w  = k + w
    
  }
  CL_Avg_B[n] <- sum(All_stats_batting)
  All_stats_batting <- c()
  w = 0
}

#print(CL_Avg_B)
```

```{r}
CL_OBP <- (CL_Avg_B[[1]]+CL_Avg_B[[3]]+CL_Avg_B[[4]])/(CL_Avg_B[[2]]+CL_Avg_B[[3]]+CL_Avg_B[[4]]+CL_Avg_B[[5]])
CL_SLG <- (CL_Avg_B[[6]])/(CL_Avg_B[[2]])
CL_wOBA <- ((.69)*(CL_Avg_B[["BB"]])+(.72)*(CL_Avg_B[["HBP"]])+(.89)*(CL_Avg_B[["X1B"]])+(1.27)*(CL_Avg_B[["2B"]])+(1.62)*(CL_Avg_B[["3B"]])+(2.10)*(CL_Avg_B[["HR"]]))/(CL_Avg_B[["AB"]]+CL_Avg_B[["BB"]]+CL_Avg_B[["SF"]]+CL_Avg_B[["HBP"]])
CL_PA <- CL_Avg_B[["AB"]]+CL_Avg_B[["BB"]]+CL_Avg_B[["HBP"]]+CL_Avg_B[["SF"]] 
CL_R_per_PA <- CL_Avg_B[["R"]]/CL_PA
CL_wRC <- CL_Avg_B[["R"]]
CL_wRC_per_PA <- CL_R_per_PA

#League adjustment for Batting Runs
ABF <- ((0.47 * CL_Avg_B[["X1B"]]) + (0.38*CL_Avg_B[["2B"]]) + (0.55*CL_Avg_B[["3B"]]) + (0.93*CL_Avg_B[["HR"]]) +(0.33*(CL_Avg_B[["BB"]]+CL_Avg_B[["HBP"]]))/(CL_Avg_B[["AB"]]-CL_Avg_B[["H"]]))

```

```{r}
bat_stats2 <- function(y){
  x <- Batting[[y]]
  #OPS+
  x$OPS_plus <- round((100)*(((x$OB/CL_OBP)+(x$SLG/CL_SLG))-1),2)
  #weighted On-Base Average
  x$wOBA <- round(((.69)*(x$BB)+(.72)*(x$HBP)+(.89)*(x$X1B)+(1.27)*(x[["2B"]])+(1.62)*(x[["3B"]])+(2.10)*(x$HR))/(x$AB+x$BB+x$SF+x$HBP),3)
  #Batting Runs
  x$Bat_Runs <- round(((0.47*x$X1B) + (0.38*x[["2B"]]) + (0.55*x[["3B"]]) + (0.93*x$HR) + ((0.33)*x$BB + x$HBP)/(ABF *(x$AB - x$H))),2)
  #weighted Runs Above Average
  x$wRAA <- round(((x$wOBA-CL_wOBA)/(1.157))*(x$PA),3)
  #weighted Runs Created
  x$wRC <- round(x$wRAA + (x$PA)*((CL_Avg_B[["R"]])/(CL_PA)),2)
  x$wRC_plus <- round((100)*((((x$wRAA/x$PA)+CL_R_per_PA)+(CL_R_per_PA-(p_factor[[y]])*(CL_R_per_PA)))/(CL_wRC_per_PA)),2)
  x
}
```

```{r lapply}
Batting <- teams %>%
  lapply(bat_stats2)
names(Batting) <- teams
```

---

---

# Top Hitters in the Crossroads League

After compiling advanced batting statistics for each team, we chose to use weighted runs created plus (wRC+) in order to evaluate a hitter's overall contribution to the team. This statistic takes park factor into account, as well as adjusting based on a player's performance in comparison to the league average.

```{r echo=FALSE}
Best_Players <- data.frame(wRC_plus=numeric())
```

```{r best_players_calc}
#counter for teams
y = 1


for (x in Batting){
  
  wRC_plus <- c()
  wRC_name <- c()
  wRC_team <- c()
  PA <- c()
  H <- c()
  HR <- c()
  RBI <- c()
  AVG <- c()
  H <- x[["H"]]
  HR <- x[["HR"]]
  RBI <- x[["RBI"]]
  AVG <- x[["AVG"]]
  wRC_plus <- x[["wRC_plus"]]
  wRC_name <- x[["Batting"]]
  PA <- x[["PA"]]
  team <- teams[y]
  for (z in wRC_name){
  
  wRC_team <- c(wRC_team, team[z])      
  
  
  }
  
  new.Best_Players <- data.frame(team, wRC_name,  wRC_plus,  PA, H, HR, RBI, AVG)
  Best_Players <- rbind(Best_Players, new.Best_Players)
  
  y = y + 1
}
```

```{r remove_values}
#remove NaN values
Best_Players <- Best_Players[complete.cases(Best_Players), ]

#remove Total and Opponents values
Best_Players <- Best_Players[ !(Best_Players$wRC_name %in% c("Opponents:", "Total:")), ]

Best_Players <- Best_Players[order(-Best_Players$wRC_plus),]

#require players to have at least 20 plate appearances
Best_Players <- Best_Players[!(Best_Players$PA < 20),]

```

The following is a list of the top hitters in the Crossroads League in 2019 based on wRC+.

```{r all_hitters_table}
num_players <- NROW(Best_Players)
index <- seq(1,num_players,1)
rownames(Best_Players) <- index

datatable(Best_Players, options = list(pageLength = 10), colnames = c('Rank' = 1, 'Team' = 2, 'Player' = 3, 'wRC+' = 4))
```

---

---

```{r load pitchstats}
load("Data/stats_by_type.Rdata")
Pitching <- baseball[["Pitching"]]
teams <- names(Pitching)
```

```{r pitch variables}
FIP_stats <- c("ER", "HR", "SO", "HBP", "BB", "IP","ERA")
CL_Avg_P <- c()
All_stats_pitching <- c()
```

```{r CL pitch stats}
w = 0

for (n in seq(length(FIP_stats)-1)){
  
  for (j in teams){
    
    P <- read.csv(paste("Data/Pitching/",j,".csv",sep=""))
    
    
    for (k in seq(length(P[[FIP_stats[[n]]]]))){ 
      
      All_stats_pitching[w+k] <- P[[FIP_stats[[n]]]][k]
      #print(All_stats)
    }
    
    w  = k + w
    
  }
  CL_Avg_P[n] <- mean(All_stats_pitching)
  All_stats_pitching <- c()
  w = 0
}

CL_ERA <- (CL_Avg_P[[1]]*9)/(CL_Avg_P[[6]])
#print(CL_ERA)
```

```{r FIP constant}
FIP_CL_constant = CL_ERA - (((13*CL_Avg_P[[2]])+(3*(CL_Avg_P[[5]]+CL_Avg_P[[4]]))-(2*CL_Avg_P[[3]]))/(CL_Avg_P[[6]]))
#print(FIP_CL_constant)
```

```{r FIP loop}
pitch_stats <- function(y){
  x <- Pitching[[y]]
  FIP_team <- ((13*x$HR)+3*(x$BB+x$HBP)-(2*x$SO))/(x$IP)
  x$FIP <- round(FIP_team + FIP_CL_constant, 2)
  x$PR <- round((x$IP*(CL_ERA - x$ERA))/9, 2)
  x
}

Pitching <- teams %>%
  lapply(pitch_stats)
names(Pitching) <- teams
```

```{r best_players_calc pitch}
Best_Pitchers <- data.frame(FIP=numeric())

#counter for teams
y = 1


for (x in Pitching){
  
  FIP <- c()
  FIP_name <- c()
  FIP_team <- c()
  IP <- c()
  SO <- c()
  ERA <- c()
  SO <- x[["SO"]]
  FIP <- x[["FIP"]]
  FIP_name <- x[["Pitching"]]
  IP <- x[["IP"]]
  ERA <- x[["ERA"]]
  team <- teams[y]
  for (z in FIP_name){
  
  FIP_team <- c(FIP_team, team[z])      
  
  
  }
  
  new.Best_Pitchers <- data.frame(team, FIP_name,  FIP,  IP, SO, ERA)
  Best_Pitchers <- rbind(Best_Pitchers, new.Best_Pitchers)
  
  y = y + 1
}
```

```{r remove_values pitchers}
#remove NaN values
Best_Pitchers <- Best_Pitchers[complete.cases(Best_Pitchers), ]

#remove Total and Opponents values
Best_Pitchers <- Best_Pitchers[ !(Best_Pitchers$FIP_name %in% c("Opponents:", "Total:")), ]

#require players to have at least 25 innings pitched
Best_Pitchers <- Best_Pitchers[!(Best_Pitchers$IP < 25),]

Best_Pitchers <- Best_Pitchers[order(Best_Pitchers$FIP),]

```

# Top Pitchers in the Crossroads League

We chose to measure pitchers using Fielding Indepent Pitching (FIP). This metric attempts to isolate the performance of the pitcher, stripping away the influence of the defense as well as luck. Similar to the wRC+ stat for hitters, FIP also adjusts based on a league constant. The scale for FIP is similar to ERA; a lower FIP is better.

The following is a list of the top pitchers in the Crossroads League in 2019 based on FIP (minimum 25 IP).

```{r all_pitchers_table}
num_players <- NROW(Best_Pitchers)
index <- seq(1,num_players,1)
rownames(Best_Pitchers) <- index

datatable(Best_Pitchers, options = list(pageLength = 10), colnames = c('Rank' = 1, 'Team' = 2, 'Player' = 3, 'K' = 6))
```

```{r schedules}
library(rvest)
library(stringr)
library(tidyverse)

#2019 schedules

load("Data/stats_by_type.Rdata")
Batting <- baseball[["Batting"]]
Pitching <- baseball[["Pitching"]]
Fielding <- baseball[["Fielding"]]
teams <- names(Batting)

#General info
teams <- c("BC","GOC","GRC","HU","INWU","MAR","MVNU","SAU","SFIN","TAYL")
names(teams) <- c(1629, 1678, 1679, 1688, 1694, 1717, 1736, 1780, 1805, 1784) 

#get schedules from dakstats
schedule_urls <- paste0("http://www.dakstats.com/WebSync/Pages/Team/TeamSchedule.aspx?association=10&sg=MBA&sea=NAIMBA_2019&team=",names(teams))
#View(schedule_urls)

#function to get the schedule of a team
get_schedule <- function(schedule_urls) {
  bpf <- schedule_urls %>%
    read_html() %>%
    html_nodes("table") %>%
    .[[36]] %>%
    html_table(fill = TRUE)
}

#p_effects_urls <- teams %>%
  #lapply(get_schedule)
#names(p_effects_urls) <- teams

schedules <- lapply(schedule_urls, get_schedule)
names(schedules) <- teams
#View(schedules)

#A function to clean up the tables and show only conference schedule
conference <- function(x){
  #Gets rid of rows containing only NA
  schedules[[x]] <- schedules[[x]][!apply(is.na(schedules[[x]]) | schedules[[x]] == "", 1, all),]
  #Rename columns
  names(schedules[[x]])[1] <- "Date"
  names(schedules[[x]])[2] <- "Opponent"
  names(schedules[[x]])[3] <- "Location"
  names(schedules[[x]])[4] <- "Score"
  names(schedules[[x]])[5] <- "Decision"
  #Grab only the conference games
  CL_schedule <- schedules[[x]] %>%
    filter(str_detect(Opponent, "\\*"))
  #remove column with only NA at the end
  CL_schedule <- CL_schedule[-6]
  #separate score into home and away columns
  CL_schedule <- CL_schedule %>%
    separate(Score, c("Team_Score","Opp_Score"),"-")
  #output only the CL schedules
  schedules[[x]] <- CL_schedule
  schedules[[x]] <- schedules[[x]] %>%
    mutate(Team = x)
}

schedules <- lapply(teams, conference)
names(schedules) <- teams

schedules_19 <- schedules
```


```{r echo=FALSE}
#A function to write the data to csv files
writeit <- function(x) {
  write.csv(schedules[[x]], file = paste0("Data/Schedules/2019", "/", x, ".csv"))
}

#Write to csv
#lapply(teams, writeit)
```


```{r sched-20}
#2020 schedules

load("Data/stats_by_type.Rdata")
Batting <- baseball[["Batting"]]
Pitching <- baseball[["Pitching"]]
Fielding <- baseball[["Fielding"]]
teams <- names(Batting)

#General info
teams <- c("BC","GOC","GRC","HU","INWU","MAR","MVNU","SAU","SFIN","TAYL")
names(teams) <- c(1629, 1678, 1679, 1688, 1694, 1717, 1736, 1780, 1805, 1784) 

#get schedules from dakstats
schedule_urls <- paste0("http://www.dakstats.com/WebSync/Pages/Team/TeamSchedule.aspx?association=10&sg=MBA&sea=NAIMBA_2020&team=",names(teams))
#View(schedule_urls)

#function to get the schedule of a team
get_schedule <- function(schedule_urls) {
  bpf <- schedule_urls %>%
    read_html() %>%
    html_nodes("table") %>%
    .[[36]] %>%
    html_table(fill = TRUE)
}

#p_effects_urls <- teams %>%
#lapply(get_schedule)
#names(p_effects_urls) <- teams

schedules <- lapply(schedule_urls, get_schedule)
names(schedules) <- teams
#View(schedules)

#A function to clean up the tables and show only conference schedule
conference <- function(x){
  #Gets rid of rows containing only NA
  schedules[[x]] <- schedules[[x]][!apply(is.na(schedules[[x]]) | schedules[[x]] == "", 1, all),]
  #Rename columns
  names(schedules[[x]])[1] <- "Date"
  names(schedules[[x]])[2] <- "Opponent"
  names(schedules[[x]])[3] <- "Location"
  names(schedules[[x]])[4] <- "Score"
  names(schedules[[x]])[5] <- "Decision"
  #Grab only the conference games
  CL_schedule <- schedules[[x]] %>%
    filter(str_detect(Opponent, "\\*"))
  #remove column with only NA at the end
  CL_schedule <- CL_schedule[-6]
  #separate score into home and away columns
  CL_schedule <- CL_schedule %>%
    separate(Score, c("Team_Score","Opp_Score"),"-")
  #output only the CL schedules
  schedules[[x]] <- CL_schedule
  schedules[[x]] <- schedules[[x]] %>%
    mutate(Team = x)
}

schedules <- lapply(teams, conference)
names(schedules) <- teams

schedules_20 <- schedules
```


```{r echo=FALSE}
#A function to write the data to csv files
writeit <- function(x) {
  write.csv(schedules[[x]], file = paste0("Data/Schedules/2020", "/", x, ".csv"))
}

#Write to csv
#lapply(teams, writeit)
```

```{r elo_initial}
library(tidyverse)
library(elo)

#place to store each team's elo rating
ratings <- data.frame(team = teams)
ratings <- ratings %>%
  mutate(elo = 1500)
#View(ratings)

#function that adds a binary variable
#1 if home wins/away loses and 0 if home loses/away wins
add_elo_score <- function(x){
  schedules_19[[x]] <- schedules_19[[x]] %>%
    mutate(result = ifelse(grepl("W",Decision,fixed=TRUE)&grepl("H",Location,fixed=TRUE),1,ifelse(grepl("A",Location,fixed=TRUE)&grepl("L",Decision, fixed=TRUE),1,0)))
  #change opponent name to match abbreviations
  schedules_19[[x]] <- schedules_19[[x]] %>%
    mutate(Opp = ifelse(grepl("Bethel (Ind.) *",Opponent,fixed=TRUE),teams[1],ifelse(grepl("Goshen (Ind.) *",Opponent,fixed=TRUE),teams[2],ifelse(grepl("Grace (Ind.) *",Opponent,fixed=TRUE),teams[3],ifelse(grepl("Huntington (Ind.) *",Opponent,fixed=TRUE),teams[4],ifelse(grepl("Indiana Wesleyan *",Opponent,fixed=TRUE),teams[5],ifelse(grepl("Marian (Ind.) *",Opponent,fixed=TRUE),teams[6],ifelse(grepl("Mount Vernon Nazarene (Ohio) *",Opponent,fixed=TRUE),teams[7],ifelse(grepl("Spring Arbor (Mich.) *",Opponent,fixed=TRUE),teams[8],ifelse(grepl("St. Francis (Ind.) *",Opponent,fixed=TRUE),teams[9],teams[10]))))))))))
}

schedules_19 <- lapply(teams, add_elo_score)
names(schedules_19) <- teams
```

```{r end_elos}
#creating a master schedule with all games
AllGames <- do.call("rbind",schedules_19)
HomeGames <- AllGames %>%
  dplyr::filter(str_detect(Location, "H"))
#View(HomeGames)


#Fix extra innings issue
Opp <- HomeGames[["Opp_Score"]]

for (x in seq(length(Opp))){
  if (grepl("(", Opp[x], fixed = TRUE) == TRUE){
     Opp[x] <- as.integer(substr(Opp[x], 1, nchar(Opp[x])-4))
    
    }
}

HomeGames[["Opp_Score"]] <- Opp

#Add margin of victory column

HomeGames$Team_Score <- as.integer(HomeGames$Team_Score)
HomeGames$Opp_Score <- as.integer(HomeGames$Opp_Score)

HomeGames$Vict_Margin <- (HomeGames$Team_Score - HomeGames$Opp_Score)

#Way to run elo that includes margin of victory using log

#Run elo
e <- elo.run(score(HomeGames[["Team_Score"]], HomeGames[["Opp_Score"]]) ~ HomeGames$Team + HomeGames$Opp +
               k(20*log(abs(HomeGames[["Team_Score"]] - HomeGames[["Opp_Score"]]) + 1)), data = HomeGames)


#print results
f <- final.elos(e)
f <- sort(f,decreasing = TRUE)

```

```{r}
#regress elo values to the mean
for (g in 1:length(f)){

   f[g] <- f[g] * (2/3) + 500
  
}

#View(f)
```

---

---

# Upcoming Games

```{r update_web}
update_web <- function(team1, team2, game_loc){
  
  team1_elo <- f[team1]
  team2_elo <- f[team2]
  #print(team1_elo)
  #print(team2_elo)
  game_prob_team1win <- elo.prob(team1_elo, team2_elo)
  game_prob_team2win <- elo.prob(team2_elo, team1_elo)
  
  print(game_prob_team1win)
  print(game_prob_team2win)
  game_park_effect <- p_factor[game_loc]
  print(game_park_effect)
  #print(f[team1])
  elos <- c(f[team1], f[team2])
  teams <- names(elos)

  elo_data = data.frame(teams, elos)
  elo_data <- elo_data %>% 
    mutate(Game = paste0(teams[1], " vs ", teams[2])) %>%
    mutate(elo_label = paste0(teams, " has elo score of ", round(elos)))
  elo_graph <- ggplot(elo_data, aes(x = teams, y = elos, fill = teams)) +
    geom_col() +
    geom_text(aes(label = elo_label), position = position_stack(vjust = 0.5)) +
    scale_fill_brewer(palette = "Set2")
  
  #The format below will work well once you have perent chance of winning
  
  elos_probs <- c(game_prob_team1win, game_prob_team2win)
  teams_probs <- c(team1, team2)
  elo_data_probs = data.frame(teams_probs, elos_probs)
  
  elo_data_probs <- elo_data_probs %>% 
    mutate(Game = paste0(teams[1], " vs ", teams[2])) %>%
    mutate(elo_label = paste0(teams, " has win probability of ", round(elos_probs, 6)))
  
  prob_graph <- ggplot(elo_data_probs, aes(x = Game, y = elos_probs, fill = teams)) +
    geom_col() +
    geom_text(aes(label = elo_label), position = position_stack(vjust = 0.5)) +
    scale_fill_brewer(palette = "Set2")
  
  
  library(ggpubr)
  ggarrange(elo_graph, prob_graph, ncol=1, nrow=2)
  
}
```

The next game is Indiana Wesleyan vs. Grace at Grace.
```{r}
update_web("INWU", "GRC", "GRC")
```

---

---

# Simulation

Due to the COVID-19 outbreak in 2020, we were unable to compare our predictions to the actual outcome of games. Instead, we have created a tournament simulation that uses the 2019 elo ratings along with the 2020 preseason coaches' poll to simulate what the season might have been like. Since this simulation uses random numbers, each time you refresh or check the website, the results will be slightly different.

```{r simulation_for_results}
#coaches poll
ranks <- c(1:10)
team_ranks <- c("HU", "MVNU", "INWU", "TAYL", "MAR", "SAU", "BC", "GOC", "GRC", "SFIN")
adjustment <- c()

#adjustment for coaches poll (function can be changed)
for (counter in ranks){
  
  adjust_value <- counter*-1
  adjustment[counter] <- adjust_value
  #print(adjustment)
}



coaches_poll <- data.frame(ranks, adjustment, row.names = team_ranks)


#custom probability function to add randomness for tournament
#custom_prob <- function(elo.A, elo.B, adjust.A, adjust.B)
#{
 # 1/(1 + 10^(((elo.B + adjust.B) - (elo.A + ifelse(elo.A > 1500, adjust.A / 2, adjust.A)))/400.0))
#}


#Create dataframe of all matchups (add loop to have teams play multiple times)
Home_tourney_team <- c()
Away_tourney_team <- c()
tourney_matchups <- data.frame(Home_tourney_team, Away_tourney_team)


for (i in 1:(length(teams)-1)){
  for (j in 1:(length(teams)-i)){
    matchup <- data.frame(
      Home_tourney_team = teams[i],
      Away_tourney_team = teams[i+j]
    )
    #names(matchup) <- c("Home", "Away")
    tourney_matchups <- rbind(tourney_matchups, matchup)
    
  }
  
}

tourney_matchups_randomized <- tourney_matchups[sample(nrow(tourney_matchups)),]


tourney_probs <- c()
winners <- c()
Home_elos <- c()
Away_elos <- c()
add_elo_score <- function(x){
  schedules_20[[x]] <- schedules_20[[x]] %>%
    mutate(result = ifelse(grepl("W",Decision,fixed=TRUE)&grepl("H",Location,fixed=TRUE),1,ifelse(grepl("A",Location,fixed=TRUE)&grepl("L",Decision, fixed=TRUE),1,0)))
  #change opponent name to match abbreviations
  schedules_20[[x]] <- schedules_20[[x]] %>%
    mutate(Opp = ifelse(grepl("Bethel (Ind.) *",Opponent,fixed=TRUE),teams[1],ifelse(grepl("Goshen (Ind.) *",Opponent,fixed=TRUE),teams[2],ifelse(grepl("Grace (Ind.) *",Opponent,fixed=TRUE),teams[3],ifelse(grepl("Huntington (Ind.) *",Opponent,fixed=TRUE),teams[4],ifelse(grepl("Indiana Wesleyan *",Opponent,fixed=TRUE),teams[5],ifelse(grepl("Marian (Ind.) *",Opponent,fixed=TRUE),teams[6],ifelse(grepl("Mount Vernon Nazarene (Ohio) *",Opponent,fixed=TRUE),teams[7],ifelse(grepl("Spring Arbor (Mich.) *",Opponent,fixed=TRUE),teams[8],ifelse(grepl("St. Francis (Ind.) *",Opponent,fixed=TRUE),teams[9],teams[10]))))))))))
}

schedules_20 <- lapply(teams, add_elo_score)
names(schedules_20) <- teams

#creating a master schedule with all games
AllGames_20 <- do.call("rbind",schedules_20)
HomeGames_20 <- AllGames_20 %>%
  filter(str_detect(Location, "H"))
#View(HomeGames_20)


#Fix extra innings issue
Opp <- HomeGames_20[["Opp_Score"]]

for (x in seq(length(Opp))){
  if (grepl("(", Opp[x], fixed = TRUE) == TRUE){
    Opp[x] <- as.integer(substr(Opp[x], 1, nchar(Opp[x])-4))
    
  }
}

HomeGames_20[["Opp_Score"]] <- Opp

#Add margin of victory column

HomeGames_20$Team_Score <- as.integer(HomeGames_20$Team_Score)
HomeGames_20$Opp_Score <- as.integer(HomeGames_20$Opp_Score)

HomeGames_20$Vict_Margin <- (HomeGames_20$Team_Score - HomeGames_20$Opp_Score)

results <- elo.run(score(Team_Score, Opp_Score) ~ adjust(Team,24) + Opp +
                     k(20*log(abs(Team_Score - Opp_Score) + 1)), data = HomeGames_20, initial.elos = f, na.action = na.omit)
#View(results)

#create vector with results after 2020 and add Spring Arbor who played no games back
After_2020_games_elos <- append(final.elos(results), f["SAU"])
#Use After_2020_games_elos for elos after 2020 games or f for elos after 2019
tourney_elos <- After_2020_games_elos
#need a loop with win probabilities for each game and adjustment of elo values
for (games in 1:nrow(tourney_matchups_randomized)){
  
  
  #get teams for game
  Home_tourney_temp <- tourney_matchups_randomized[games,1]
  Away_tourney_temp <- tourney_matchups_randomized[games,2]
  #get and adjust elo values
  Home_tourney_temp_elo <- tourney_elos[toString(Home_tourney_temp)] + rnorm(1, mean=1+coaches_poll[Home_tourney_temp,]$adjustment, sd=25)#+ coaches_poll[toString(Home_tourney_temp), "adjustment"]
  Away_tourney_temp_elo <- tourney_elos[toString(Away_tourney_temp)] + rnorm(1, mean=1+coaches_poll[Away_tourney_temp,]$adjustment, sd=25)#+ coaches_poll[toString(Away_tourney_temp), "adjustment"]
  
  #add elos to vector for dataframe
  Home_elos[games] <- Home_tourney_temp_elo
  Away_elos[games] <- Away_tourney_temp_elo
  
  #get home team win probability
  game_probs <- elo.prob(Home_tourney_temp_elo, Away_tourney_temp_elo)#elo.prob(f[toString(Home_tourney_temp)], f[toString(Away_tourney_temp)])
  
  #keep track of winners
  if (game_probs > 0.5){
    
    winners[games] <- toString(Home_tourney_temp)
    win_check <- 1
    
  }
  else{
    
    winners[games] <- toString(Away_tourney_temp)
    win_check <- 0
    
  }
  
  
  
  
  Home_team_pred_score <- 10*game_probs
  Away_team_pred_score <- 10*(1-game_probs)
  
  #caculate new elo values based on predicted score
  run_elo_sim <- elo.calc(win_check, Home_tourney_temp_elo, Away_tourney_temp_elo, k(20*log(abs(Home_team_pred_score - Away_team_pred_score) + 1)))
  
  #change elo values in table f_copy
  tourney_elos[toString(Home_tourney_temp)] <- as.double(run_elo_sim["elo.A"])
  tourney_elos[toString(Away_tourney_temp)] <- as.double(run_elo_sim["elo.B"])
  
  
  #probability that home beats away
  #print(tourney_probs)
  tourney_probs <- c(tourney_probs, game_probs)

  

}

#tourney_matchups <- cbind(tourney_matchups, winners)

tourney_matchups_randomized <- cbind(tourney_matchups_randomized, Home_elos, Away_elos, tourney_probs, winners)

tourney_elos <- sort(tourney_elos,decreasing = TRUE)


tourney_results_vec <- c()
for (sim_x in (1:length(team_ranks))){
  #print(team_ranks[sim_x])
  games_won_sim <- length(winners[winners==team_ranks[sim_x]])
  games_lost_sim <- 9-games_won_sim
  #print(games_lost_sim)
  sim_record <- paste(games_won_sim,games_lost_sim, sep="-")
  tourney_results_vec[sim_x] <- sim_record
  #print(tourney_results_vec)
  
  
}

tourney_results <- data.frame(team_ranks, tourney_results_vec)
names(tourney_results)[1] <- "Team"
names(tourney_results)[2] <- "Record"
print(tourney_results)
print(tourney_elos)

#tail(tourney_elos, n=1)
```
---