
---
title: "Batting Statistics"
author: "Kara Godsey"
date: "3/10/2020"
output: html_notebook
---

```{r setup, echo=FALSE, cache=FALSE}
suppressWarnings(library(knitr))
suppressWarnings(library(rmdformats))
suppressWarnings(library(kableExtra))
suppressWarnings(library(rvest))
```

# Batting Statistics

First, we loaded the data tables that we previously created.

```{r}
load("../Data/stats_by_type.Rdata")
Batting <- baseball[["Batting"]]
teams <- names(Batting)
```

Next, we created a function to calculate advanced statistics using the base data in the tables. The statistics we chose to add to our tables include: Caught Stealing, Runs Created (Basic), OPS, POP, Total Average, Batting Average on Balls in Play (BABIP), and the Hoban Efficiency Quotient for offense.

```{r}
bat_stats <- function(y){
  x <- Batting[[y]]
  #Caught Stealing
  x$CS <- x$SBA - x$SB
  #Singles
  x$X1B <- x$H - x[,8] - x[,9] - x$HR
  #Runs Created
  x$RC <- round((x$H + x$BB)*x$TB/(x$AB + x$BB),2)
  #OPS
  x$OPS <- x$OB + x$SLG
  #POP
  x$POP <- x$OPS + x$AVG
  #Total Average
  x$TA <- round((x$TB + x$BB + x$HBP + x$SB)/(x$AB - x$H + x$SH + x$SF + x$CS + x$GDP), 3)
  #Plate Appearances
  x$PA <- x$AB+x$BB+x$HBP+x$SF
  #Batting Average on Balls in Play
  x$BABIP <- round((x$H - x$HR)/(x$AB - x$SO - x$HR + x$SF),3)
  #Hoban Efficiency Quotient - offense
  x$HEQO <- x$TB + x$R + x$RBI + x$SB + (0.5*x$BB)
  #Base Stealing Runs
  x$Steal_Runs <- round(((x$H + x$BB - x$CS)*(x$TB + (0.7 * x$SB)))/(x$AB + x$BB + x$CS),2)
  x
}
```

Then we used the `lapply` R function to apply these statistics to each of the teams, resulting in advanced statistics for every player in the league.

```{r lapplyit}
Batting <- teams %>%
  lapply(bat_stats)
names(Batting) <- teams
```

Next, we calculated league totals for the basic statistics, for the purpose of calculating more advanced individual statistics later on.

```{r}
batting_stats <- c("H","AB","BB","HBP","SF","TB", "2B","3B","HR","X1B","R")
CL_Avg_B <- c()
All_stats_batting <- c()
w = 0

for (n in batting_stats){
  
  for (j in teams){
    
    P <- Batting[[j]]#as.data.frame(read.csv(paste("Data/Batting/",j,".csv",sep="")))
    
    col_num =  which( colnames(P)==n )
    for (k in seq(length(P[[n]]))){ 
      
      All_stats_batting[w+k] <- P[,col_num][k]
      #print(All_stats)
    }
    
    w  = k + w
    
  }
  CL_Avg_B[n] <- sum(All_stats_batting)
  All_stats_batting <- c()
  w = 0
}

print(CL_Avg_B)
```

Then, we calculated advanced statistics for the Crossroads League as a whole. These included: On-Base Percentage (OBP), Slugging Percentage (SLG), weighted On-Base Percentage (wOBA), Plate Appearances (PA), Runs per Plate Appearance (R/PA), weighted Runs Created (wRC), and weighted Runs Created per Plate Appearance (wRC/PA). Additionally, we calculated ABF, which is a constant that is necessary to caluclate batting runs later on.

```{r}
CL_OBP <- (CL_Avg_B[[1]]+CL_Avg_B[[3]]+CL_Avg_B[[4]])/(CL_Avg_B[[2]]+CL_Avg_B[[3]]+CL_Avg_B[[4]]+CL_Avg_B[[5]])
CL_SLG <- (CL_Avg_B[[6]])/(CL_Avg_B[[2]])
CL_wOBA <- ((.69)*(CL_Avg_B[["BB"]])+(.72)*(CL_Avg_B[["HBP"]])+(.89)*(CL_Avg_B[["X1B"]])+(1.27)*(CL_Avg_B[["2B"]])+(1.62)*(CL_Avg_B[["3B"]])+(2.10)*(CL_Avg_B[["HR"]]))/(CL_Avg_B[["AB"]]+CL_Avg_B[["BB"]]+CL_Avg_B[["SF"]]+CL_Avg_B[["HBP"]])
CL_PA <- CL_Avg_B[["AB"]]+CL_Avg_B[["BB"]]+CL_Avg_B[["HBP"]]+CL_Avg_B[["SF"]] 
CL_R_per_PA <- CL_Avg_B[["R"]]/CL_PA
CL_wRC <- CL_Avg_B[["R"]]
CL_wRC_per_PA <- CL_R_per_PA

#League adjustment for Batting Runs
ABF <- ((0.47 * CL_Avg_B[["X1B"]]) + (0.38*CL_Avg_B[["2B"]]) + (0.55*CL_Avg_B[["3B"]]) + (0.93*CL_Avg_B[["HR"]]) +(0.33*(CL_Avg_B[["BB"]]+CL_Avg_B[["HBP"]]))/(CL_Avg_B[["AB"]]-CL_Avg_B[["H"]]))

```

Now that we have these league statistics, we can add more advanced statistics to our tables. The statistics we added were: OPS+, wOBA, Batting Runs, weighted Runs Above Average (wRAA), weighted Runs Created (wRC), and weighted Runs Created plus (wRC+). All of these statistics adjust the individual player's performance to reflect their contribution in comparison to other players in the league.  

```{r}
bat_stats2 <- function(y){
  x <- Batting[[y]]
  #OPS+
  x$OPS_plus <- round((100)*(((x$OB/CL_OBP)+(x$SLG/CL_SLG))-1),2)
  #weighted On-Base Average
  x$wOBA <- round(((.69)*(x$BB)+(.72)*(x$HBP)+(.89)*(x$X1B)+(1.27)*(x[["2B"]])+(1.62)*(x[["3B"]])+(2.10)*(x$HR))/(x$AB+x$BB+x$SF+x$HBP),3)
  #Batting Runs
  x$Bat_Runs <- round(((0.47*x$X1B) + (0.38*x[["2B"]]) + (0.55*x[["3B"]]) + (0.93*x$HR) + ((0.33)*x$BB + x$HBP)/(ABF *(x$AB - x$H))),2)
  #weighted Runs Above Average
  x$wRAA <- round(((x$wOBA-CL_wOBA)/(1.157))*(x$PA),3)
  #weighted Runs Created
  x$wRC <- round(x$wRAA + (x$PA)*((CL_Avg_B[["R"]])/(CL_PA)),2)
  x$wRC_plus <- round((100)*((((x$wRAA/x$PA)+CL_R_per_PA)+(CL_R_per_PA-(p_factor[[y]])*(CL_R_per_PA)))/(CL_wRC_per_PA)),2)
  x
}
```

Finally, we used the `lapply` R function again to apply these statistics to all the teams in the Crossroads League.

```{r lapply}
Batting <- teams %>%
  lapply(bat_stats2)
names(Batting) <- teams
```

Here is an example of the resulting tables that we produced.

```{r echo=FALSE}
kable(Batting[["GRC"]]) %>%
  kable_styling(bootstrap_options = c("striped", "bordered", "condensed"), full_width = FALSE)
```

---