---
title: "Scraping the Website"
author: "Ryan Johnson"
output:
  html_document:
    toc: true
    numbered_sections: true
    theme: united
    highlight: tango
---

For this project we used the NAIA statistics posted on the website [dakstats](http://www.dakstats.com/websync/Pages/Association.aspx?association=10).  They have batting, pitching, and fielding statitistics for each baseball team in the Crossroads league, of which Grace College is a member.

We found the `rvest` R package to be very useful for this project.
```{r rvest}
suppressWarnings(library(rvest))
```

Dakstats assigns each team an identification number, and uses this number in their webpage urls.  We've created 2-4 letter abreviations for each team, and named them by their ID numbers.
```{r teams-and-urls}
teams <- c("BC","GOC","GRC","HU","INWU","MAR","MVNU","SAU","SFIN","TAYL")
names(teams) <- c(1629, 1678, 1679, 1688, 1694, 1717, 1736, 1780, 1805, 1784)
urls <- paste0("http://www.dakstats.com/WebSync/Pages/Team/IndividualStats.aspx?association=10&sg=MBA&conference=NAIMBA1_CROSS&team=",names(teams),"&sea=NAIMBA_2019")
```

Then we created functions to retrieve only the batting, pitching, and fielding tables, which we turned into data frames in R.

```{r retrieval-functions}
get_batting <- function(url) {
  bpf <- url %>%
    read_html() %>%
    html_nodes("table") %>%
    .[[37]] %>%
    html_table(fill = TRUE)
}

get_pitching <- function(url) {
  bpf <- url %>%
    read_html() %>%
    html_nodes("table") %>%
    .[[38]] %>%
    html_table(fill = TRUE)
}

get_fielding <- function(url) {
  bpf <- url %>%
    read_html() %>%
    html_nodes("table") %>%
    .[[39]] %>%
    html_table(fill = TRUE)
}
```

We used the `lapply` R function with these functions to get our data.

```{r lapply-it}
batting <- lapply(urls, get_batting)
names(batting) <- teams
pitching <- lapply(urls, get_pitching)
names(pitching) <- teams
fielding <- lapply(urls, get_fielding)
names(fielding) <- teams
```
And put all of our data into one list.
```{r}
baseball <- list(batting, pitching, fielding)
names(baseball) <- c("Batting", "Pitching", "Fielding")
```
Then we wrote a function to write all of the tables into csv files,
```{r writeit}
#A function to write the data to csv files
writeit <- function(x) {
  #Batting
  write.csv(batting[[x]], file = paste0("../Data/Batting", "/", x, ".csv"))
  #Pitching
  write.csv(pitching[[x]], file = paste0("../Data/Pitching", "/", x, ".csv"))
  #Fielding
  write.csv(fielding[[x]], file = paste0("../Data/Fielding", "/", x, ".csv"))
}
```
and used `lapply` to write the csv's all at once.
```{r write-to-csv}
invisible(lapply(teams, writeit))
```
Lastly, we also saved our data to an Rdata file for our own convenience.
```{r saved}
save(baseball, file = "../Data/stats_by_type.Rdata")
```

