---
title: "Tables"
author: "Kara Godsey"
date: "3/9/2020"
output: html_notebook
---

```{r setup, echo=FALSE, cache=FALSE}
suppressWarnings(library(knitr))
suppressWarnings(library(rmdformats))
suppressWarnings(library(kableExtra))
```

# Top Hitters in the Crossroads League

After compiling advanced batting statistics for each team, we chose to use weighted runs created plus (wRC+) in order to evaluate a hitter's overall contribution to the team. This statistic takes the park factor into account. Additionally, it adjusts based on a player's performance in comparison to the league average.

```{r echo=FALSE}
Best_Players <- data.frame(wRC_plus=numeric())
```

## The Process

In order to determine the top hitters in the entire league, we needed to merge all the data into one master dataframe. In order to do this, we created a for loop to collect the players and their stats from each team's dataframes.

```{r}
#counter for teams
y = 1


for (x in Batting){
  
  wRC_plus <- c()
  wRC_name <- c()
  wRC_team <- c()
  PA <- c()
  wRC_plus <- x[["wRC_plus"]]
  wRC_name <- x[["Batting"]]
  PA <- x[["PA"]]
  team <- teams[y]
  for (z in wRC_name){
  
  wRC_team <- c(wRC_team, team[z])      
  
  
  }
  
  new.Best_Players <- data.frame(team, wRC_name,  wRC_plus,  PA)
  Best_Players <- rbind(Best_Players, new.Best_Players)
  
  y = y + 1
}
```

Next, we cleaned up the data a bit by removing non-zero values and the rows containing team totals and opponent totals. Finally, we implemented a requirement of at least 20 plate appearances.

```{r}
#remove NaN values
Best_Players <- Best_Players[complete.cases(Best_Players), ]

#remove Total and Opponents values
Best_Players <- Best_Players[ !(Best_Players$wRC_name %in% c("Opponents:", "Total:")), ]

Best_Players <- Best_Players[order(-Best_Players$wRC_plus),]

#require players to have at least 20 plate appearances
Best_Players <- Best_Players[!(Best_Players$PA < 20),]

```

## Top 10 Hitters

The following is a list of the top 10 hitters in the Crossroads League in 2019.

```{r echo=FALSE}
Best_Players$wRC_plus <- round(Best_Players$wRC_plus, 2)
kable(Best_Players[1:10, 1:4], row.names = FALSE, col.names=c("Team", "Player", "wRC+", "PA")) %>%
  kable_styling(bootstrap_options = c("striped", "bordered"), full_width = FALSE) %>%
    footnote(general = "*min. 20 plate appearances", general_title = "")
```

---