---
title: "Tables"
author: "Kara Godsey"
date: "3/9/2020"
output: html_notebook
---

```{r setup, echo=FALSE, cache=FALSE}
suppressWarnings(library(knitr))
suppressWarnings(library(rmdformats))
suppressWarnings(library(kableExtra))
```

# Top Players in the Crossroads League

After compiling advanced batting statistics for each team, we chose to use weighted runs created plus (wRC+) in order to evaluate a hitter's overall contribution to the team. This statistic takes the park factor into account. Additionally, it adjusts based on a player's performance in comparison to the league average.

The following is a list of the top 10 hitters in the Crossroads League in 2019.

```{r echo=FALSE}
Best_Players$wRC_plus <- round(Best_Players$wRC_plus, 2)
kable(Best_Players[1:10, 1:4], row.names = FALSE, col.names=c("Team", "Player", "wRC+", "PA")) %>%
  kable_styling(bootstrap_options = c("striped", "bordered"), full_width = FALSE) %>%
    footnote(general = "*min. 20 plate appearances", general_title = "")
```